---
title: "Research Proposal"
subtitle: "Measuring Food Waste in Prince George Restaurant:\ Volume, Model, and Effects"
author: "Akihiko Mori, NRES UNBC"
date: "January 01, 2023"
output: 
  beamer_presentation:
    slide_level: 2
    toc: true
    theme: "Berkeley"
    colortheme: "whale"
    fonttheme: "structurebold"
    incremental: false
header-includes: 
- \definecolor{unbc}{HTML}{035642}
- \setbeamercolor{structure}{fg=unbc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

## Introduction

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

# Objectives

## Objectives

- Bullet 1
- Bullet 2
- Bullet 3

# Literature Reviews

## Literature Reviews

```{r cars, echo = TRUE}
summary(cars)
```

# Methods

## Methods

```{r pressure}
plot(pressure)
```

# Expected Results

## Expected Results

- AAA
- 
